A Fast and Tight Heuristic for A* in Road Networks

### 2021-08-09 paper/ch_potentials.tex
remove pagebreaks
remove line numbers
prepare for arxiv
add some very dirty writing on many-to-{one,many} lazy rphast

@@

\newpage

@@

\begin{algorithm2e}
\KwData{$B[x]$: tentative distance from $x$ to $t$ as computed by Algorithm~\ref{algo:ch-backward}}
\KwData{$P[x]$: memoized potential at $x$, $\bot$ initially}
\SetKwFunction{Pot}{Pot}
\SetKwProg{Fn}{Function}{:}{}
\Fn{\Pot{$x$}}{
	\If{$P[x] = \bot$}{
		$P[x]\leftarrow B[x]$\;
		\For{all up edges $xy$ in $G^+_\ell$}{
                        $P[x]\leftarrow\min\{P[x],w_\ell(xy)+\Pot(y)\}$\;
		}
	}
	\Return{$P[x]$}\;
}
\caption{CH-Potentials Algorithm}
\label{algo:pot}
\end{algorithm2e}

@@

@@

\pagebreak

>>>>

\nolinenumbers
\hideLIPIcs

@@

@@

\begin{algorithm2e}
\KwData{$B[x]$: tentative distance from $x$ to $t$ as computed by Algorithm~\ref{algo:ch-backward}}
\KwData{$D[x]$: memoized distance final from $x$ to $t$, $\bot$ initially}
\SetKwFunction{Dist}{ComputeAndMemoizeDist}
\SetKwProg{Fn}{Function}{:}{}
\Fn{\Dist{$x$}}{
	\If{$D[x] = \bot$}{
		$D[x]\leftarrow B[x]$\;
		\For{all up edges $xy$ in $G^+_\ell$}{
                        $D[x]\leftarrow\min\{D[x],w_\ell(xy)+\Dist(y)\}$\;
		}
	}
	\Return{$P[x]$}\;
}
\caption{Lazy RPHAST Algorithm}
\label{algo:pot}
\end{algorithm2e}

@@

\subsection{CCH-Potentials}

\begin{algorithm2e}
\KwData{$B[x]$: tentative distance from $x$ to $t$ as computed by Algorithm~\ref{algo:ch-backward}}
\KwData{$D[x]$: memoized distance final from $x$ to $t$, $\bot$ initially}
\KwData{$P[x]$: parent node of $x$ in the elimination tree}
\SetKwFunction{Dist}{ComputeAndMemoizeDist}
\SetKwProg{Fn}{Function}{:}{}
\Fn{\Dist{$x$}}{
	\If{$D[x] = \bot$}{
		\If{$D[P[x]] = \bot$}{
			$\Dist(P[x])$\;
		}
		$D[x]\leftarrow B[x]$\;
		\For{all up edges $xy$ in $G^+_\ell$}{
                        $D[x]\leftarrow\min\{D[x],w_\ell(xy)+D[y]\}$\;
		}
	}
	\Return{$P[x]$}\;
}
\caption{Elimination Tree Lazy RPHAST Algorithm}
\label{algo:cch_pot}
\end{algorithm2e}

Algorithm~\ref{algo:pot} can be applied to CCH without any modifications.
However, similarly to the elimination tree query algorithm, we can also exploit the elimination tree for the Lazy RPHAST algorithm.
Computing the distance $D[x]$ requires knowing the final distance of all upward neighbors.
In Algorithm~\ref{algo:pot} we have to compute these distances recursively.
In~\cite{}, it was proven that the set of the ancestors in the elimination tree of a node $x$ equals the CH search space of $x$. % TODO cite
Since all upward neighbors of $x$ are by definition included in the search space, they are also ancestors in the elimination tree.
If the distance of the parent in the elimination tree has already been computed, the distances of all other upward neighbors have also been computed.
Thus, it is sufficient to recursively (or iteratively) ensure that the distance of the parent in the elimination tree has been computed.
As soon as a node with final distance is encountered, all ancestors are known to have their final distance.
This allows for the more efficient implementation depicted in Algorithm~\ref{algo:cch_pot}.

\subsection{Many-to-Many with Lazy RPHAST}

In this section, we discuss how the Lazy RPHAST schema can be applied in a Many-to-Many context.
To the best of our knowledge, the fastest way to compute a full distance table between a set of sources and a set of targets is RPHAST utilizing SSE instructions to process several sources at once.
First, the selection phase is performed for the full target set.
Second, a query is performed for each source which yields the distances from this source to all targets (RPHAST).
This second step can be improved by processing $k$ (typically 16) sources at once using SIMD vector instructions (SSE RPHAST).

Since our algorithm is a variation RPHAST, we can also apply it to the Many-to-Many problem.
% TODO explain
However, a naive approach would explore the upward search space of all the source over and over again.
This naive approach would not be competitive.
Luckily, we can do better by using a different One-to-Many CH variant -- the Bucket-CH algorithm -- to select all targets at once.
The Bucket CH selection phase works as follows:
For each node $x$, we maintain a list $B[x]$ of tuples $(t_i, dist(x,t_i))$, also called a \emph{bucket}.
These buckets can be populated by running a backward CH search from each $t_i$.
An even faster way to compute the buckets is to initialize a backward CH search with all nodes at once, order the queue by node rank and use the buckets as labels.
A bucket can be propagated along an edge by increasing each entry by the edge weight, and merged with another tentative bucket with a parallel sweep over both buckets.
% TODO way too dense
Once the buckets have been computed, we can compute distances to all targets once with Algorithm~\ref{algo:pot} (or~ Algorithm~\ref{algo:cch_pot}).
The only difference is that $D[x]$ is now an array of distances to each $t_i$.
We evaluate this algorithm in Section~\ref{}.

@@
